
\section{Gaussian Processes (GPs)}
\textbf{Definition}: GP is a distribution over functions $f(\mathbf{x}) \sim
\mathcal{GP}(m(\mathbf{x}), k(\mathbf{x}, \mathbf{x}'))$, where $m(\cdot)$ is
mean function (often 0), $k(\cdot,\cdot)$ is kernel (covariance function).

\textbf{Key Kernels}:
\begin{itemize}[leftmargin=2mm]
  \item RBF: $k(\mathbf{x}, \mathbf{x}') = \sigma_f^2
    \exp\left(-\frac{\|\mathbf{x} - \mathbf{x}'\|^2}{2\ell^2}\right)$
    (lengthscale
    $\ell$, variance $\sigma_f^2$).
  \item Linear: $k(\mathbf{x}, \mathbf{x}') = \mathbf{x}^T \mathbf{x}' + c$.
  \item Matérn: $k(r) = \frac{2^{1-\nu}}{\Gamma(\nu)} \left(\frac{\sqrt{2\nu}
    r}{\ell}\right)^\nu K_\nu\left(\frac{\sqrt{2\nu} r}{\ell}\right)$ ($\nu=3/2$
    or $5/2$ for smoothness).
\end{itemize}

\textbf{GP Regression (Noisy Observations)}: $y = f(\mathbf{x}) + \epsilon$,
$\epsilon \sim \mathcal{N}(0, \sigma_n^2)$. Training data: $\mathbf{X},
\mathbf{y}$.

Prior: $\mathbf{f} \sim \mathcal{N}(\mathbf{0}, \mathbf{K})$, where $K_{ij} =
k(x_i, x_j)$.

Posterior predictive: For test points $\mathbf{X}_*$, $\mathbf{f}_* |
\mathbf{y} \sim \mathcal{N}(\bar{\mathbf{f}}_*, \cov(\mathbf{f}_*))$,
\begin{align*}
  \bar{\mathbf{f}}_* &= \mathbf{K}_{*\mathbf{X}} (\mathbf{K}_{\mathbf{XX}} +
  \sigma_n^2 \mathbf{I})^{-1} \mathbf{y}, \\
  \cov(\mathbf{f}_*) &= \mathbf{K}_{**} - \mathbf{K}_{*\mathbf{X}}
  (\mathbf{K}_{\mathbf{XX}} + \sigma_n^2 \mathbf{I})^{-1}
  \mathbf{K}_{\mathbf{X}*}.
\end{align*}

\textbf{Marginal Likelihood} (for hyperparams $\theta$): $\log p(\mathbf{y} |
\mathbf{X}, \theta) = -\frac{1}{2} \mathbf{y}^T (\mathbf{K} + \sigma_n^2
\mathbf{I})^{-1} \mathbf{y} - \frac{1}{2} \log |\mathbf{K} + \sigma_n^2
\mathbf{I}| - \frac{n}{2} \log 2\pi$.

\textbf{Tricks/Exam Hints}:
\begin{itemize}[leftmargin=2mm]
  \item Cholesky decomp for inversion: Stable for positive-definite
    $\mathbf{K}$.
  \item Optimize $\theta$ via gradient descent on log-marginal (derivs w.r.t.
    $\ell, \sigma_f$).
  \item GPs for classification: Use logistic/sigmoid + Laplace approx or EP.
  \item Scalability: Sparse GPs (e.g., FITC) approx large datasets.
\end{itemize}

\section{Ensembles}
\textbf{Key Methods}:
\begin{itemize}[leftmargin=2mm]
  \item \textbf{Bagging} (Bootstrap Aggregating): Train $M$ models on bootstrap
    samples, average predictions. Reduces variance.
  \item \textbf{Random Forest}: Bagging + random feature subsets at splits.
    Importance: $I(f) = \sum_{\text{nodes}} \Delta \text{impurity} \cdot
    p(\text{node})$.
  \item \textbf{Boosting} (e.g., AdaBoost): Sequential, weight misclassified
    points. Final: $H(\mathbf{x}) = \sign\left(\sum_m \alpha_m
    h_m(\mathbf{x})\right)$, $\alpha_m = \frac{1}{2} \log \frac{1 -
    \epsilon_m}{\epsilon_m}$.
  \item \textbf{Gradient Boosting}: Minimize loss $L = \sum_i l(y_i,
    F(\mathbf{x}_i))$, update $F_m = F_{m-1} + \nu h_m$, where $h_m$ fits
    pseudo-residuals $r_{im} = -\frac{\partial l}{\partial
    F_{m-1}(\mathbf{x}_i)}$.
\end{itemize}

\textbf{Bias-Variance Tradeoff}: Ensembles reduce variance (bagging) or bias
(boosting). Error bound for AdaBoost: $\epsilon \leq 2^M \prod_m
\sqrt{\epsilon_m (1 - \epsilon_m)}$.

\textbf{Tricks/Exam Hints}:
\begin{itemize}[leftmargin=2mm]
  \item Diversity: Key to ensembles; measure via correlation of base learners.
  \item Overfitting: Boosting can overfit; use early stopping or shrinkage $\nu
    < 1$.
  \item For proofs: Derive exponential loss minimization for AdaBoost weights.
\end{itemize}

